Discrete Quantum Theories and Computing

### 2019-01-14 dissertation.tex
dissertation.lyx "Chapter 6 Further Discussion" Updating...

\begin{comment}
When people proved the original Gleason theorem, people usually exploited
the geometrical structure of real 3-dimensional Hilbert space \cite{gleason1957,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
Since our finite-precision extension of the Gleason theorem only applies
on a class of QIVPMs, we might want to ask how to modify these geometrical
arguments to have a Gleason-type theorem for general QIVPMs. We will
further study the tensor product structure among QIVPMs which is essential
for defining product and entangled states, and serves the basis to
discuss quantum nonlocality~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and quantum computing with QIVPMs. Finally, we want to improve the
discrete quantum theories to consider QIVPMs over finite fields in
future research.
\end{comment}
Recall in previous section, we denote the maximum length of intervals
in $\mathscr{I}$ as $\left\Vert \mathscr{I}\right\Vert $. The finite-precision
extension of the Gleason theorem, Thm.~\ref{thm:Finite-precision-Gleason},
only applies on the QIVPMs of the form $f\circ\muB_{\rho}$ and states
that as $\left\Vert \mathscr{I}\right\Vert $ shrinks to $0$, the
composition $f\circ\muB_{\rho}$ would be better in the sense that
the states consistent with the composition $f\circ\muB_{\rho}$ converges
to $\rho$. In contrast, the wildest QIVPM defined in Table~\ref{tab:non-Born-QIVPM}
has the range~$\mathscr{I}_{0}$ with $\left\Vert \mathscr{I}_{0}\right\Vert =1$,
the longest possible length. This suggests that shrinking the length
$\left\Vert \mathscr{I}\right\Vert $ might regularize QIVPMs mapping
to $\mathscr{I}$ as the following question.

\begin{question}\label{question:approximation-Gleason}Given a Hilbert
space $\Hilb$ of dimension $D\geq3$, is there an $\varepsilon>0$
such that for all QIVPM $\bar{\mu}:\events\rightarrow\mathscr{I}$
satisfying $\left\Vert \mathscr{I}\right\Vert <\varepsilon$, $\bar{\mu}$
must have an non-empty core, i.e., $\coreBorn\left(\bar{\mu},\events\right)\ne\emptyset$.\end{question}

@@

To further simplify the question, for any state $\ket{\psi}\in\CP{2}$,
we can decompose the interval-valued probabilities on its projectors
$\proj{\psi}$ into the left-end and right-end on $\ket{\psi}$, i.e.,
$\bar{\mu}\left(\proj{\psi}\right)=\left[\frameL\left(\ket{\psi}\right),\frameR\left(\ket{\psi}\right)\right]$.
According to Lemma~\ref{thm:convex-3-1}, $\bar{\mu}:\events\rightarrow\mathscr{I}$
is a QIVPM if and only if the pair of functions $\frameL:\CP{2}\rightarrow\left[0,1\right]$
and $\frameR:\CP{2}\rightarrow\left[0,1\right]$ satisfying the following

@@

\left[1,1\right]-\left[\frameL\left(\ket{\psi_{0}}\right),\frameR\left(\ket{\psi_{0}}\right)\right]\subseteq\left[\frameL\left(\ket{\psi_{1}}\right),\frameR\left(\ket{\psi_{1}}\right)\right]+\left[\frameL\left(\ket{\psi_{2}}\right),\frameR\left(\ket{\psi_{2}}\right)\right]\label{eq:interval-valued-frame-function}
for any orthonormal basis $\left\{ \ket{\psi_{i}}\right\} _{i=0}^{2}$.

@@

Since $\left\Vert \mathscr{I}\right\Vert <\varepsilon$, the length

@@

In another word, the left-end function satisfies the following inequalities

@@

state $\rho=\sum_{j=1}^{N}q_{j}\proj{\Phi_{j}}$ such that 

@@

for any state $\ket{\psi}\in\CP{2}$, where $\ket{\Phi_{j}}\in\mathcal{H}$
are normalized, $q_{j}>0$, and $\sum_{j=1}^{N}q_{j}=1$. Our Question~\ref{question:approximation-Gleason}

@@

With different settings, whether there is an approximate version of
Gleason's theorem in general was asked by Sam Sanders in constructivenews~\cite{Sanders2013},
and there is no clear answer for his question as far as I understand.
To have an idea how hard this question could be, recall in Sec.~\ref{sec:fuzzy}
we state that a quantum probability space is glued by a family of
classical probability spaces. This is similar to the situation that
a manifold is glued by many local coordinates. When each small piece
has \emph{exactly} the same and positive curvature, the Killing-Hopf
theorem asserts this manifold is a sphere, but little can we say even
if the curvature has only small deviation from constant. Similar situation
might happen when approximating Gleason's theorem. Since the sum of
$\frameL$ on each ``local'' classical probability space, $\sum_{i=0}^{2}\frameL\left(\ket{\psi_{i}}\right)$,
is \emph{exactly} the same and equal to $1$, Gleason's theorem can
assert $\frameL$ can be expressed as Eq.~(\ref{eq:Gleason-for-CP2}).
However, when each local classical probability space becomes imprecise,
a general $\frameL$ might be as wild as we can imagine, and we might
say little about its global property.

@@

>>>>

\begin{comment}
When people proved the original Gleason theorem, people usually exploited
the geometrical structure of real 3-dimensional Hilbert space \cite{gleason1957,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
Since our finite-precision extension of the Gleason theorem only applies
on a class of QIVPMs, we might want to ask how to modify these geometrical
arguments to have a Gleason-type theorem for general QIVPMs. We will
further study the tensor product structure among QIVPMs which is essential
for defining product and entangled states, and serves the basis to
discuss quantum nonlocality~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and quantum computing with QIVPMs. Finally, we want to improve the
discrete quantum theories to consider QIVPMs over finite fields in
future research.
\end{comment}
As we discussed in the previous section, Thm.~\ref{thm:Finite-precision-Gleason}
only applies to the QIVPMs constructed by composing an interval map~$f$
with a conventional quantum probability measure~$\muB_{\rho}$, and
the states consistent with the composition $f\circ\muB_{\rho}$ collapses
to a single state as the maximum length of intervals in $\mathscr{I}$,
$\left\Vert \mathscr{I}\right\Vert $, shrinks to $0$. In contrast,
the globally inconsistent QIVPM defined in Table~\ref{tab:non-Born-QIVPM}
has the minimal sharp range~$\mathscr{I}_{0}$ with $\left\Vert \mathscr{I}_{0}\right\Vert =1$.
This suggests that even for general QIVPMs, shrinking the length $\left\Vert \mathscr{I}\right\Vert $
might help to regularize them. In another word, if $\left\Vert \mathscr{I}\right\Vert $
is shorter than a number, all QIVPMs map to $\mathscr{I}$ might have
an non-empty core.

\begin{question}\label{question:approximation-Gleason}Given a Hilbert
space $\Hilb$ of dimension $D\geq3$, is there an $\varepsilon>0$
such that for all QIVPM $\bar{\mu}:\events\rightarrow\mathscr{I}$
satisfying $\left\Vert \mathscr{I}\right\Vert \le\varepsilon$, $\bar{\mu}$
must have an non-empty core, i.e., $\coreBorn\left(\bar{\mu},\events\right)\ne\emptyset$.\end{question}

@@

Therefore, to study a QIVPM~$\bar{\mu}$, we can just study a pair
of functions $\frameL:\CP{2}\rightarrow\left[0,1\right]$ and $\frameR:\CP{2}\rightarrow\left[0,1\right]$
defined by $\left[\frameL\left(\ket{\psi}\right),\frameR\left(\ket{\psi}\right)\right]=\bar{\mu}\left(\proj{\psi}\right)$
for $\ket{\psi}\in\CP{2}$. According to Lemma~\ref{thm:convex-3-1},
$\bar{\mu}:\events\rightarrow\mathscr{I}$ is a QIVPM if and only
if Eqs.~(\ref{eq:QIVPM-constraints}) and
\bar{\mu}\left(\proj{\psi_{1}}+\proj{\psi_{2}}\right)\subseteq\bar{\mu}\left(\proj{\psi_{1}}\right)+\bar{\mu}\left(\proj{\psi_{2}}\right)\label{eq:interval-valued-inclusion}
holds for any orthonormal basis $\left\{ \ket{\psi_{i}}\right\} _{i=0}^{2}$.
By applying Eq.~(\ref{eq:QIVPM-complement}) on the left-hand side,
Eq.~(\ref{eq:interval-valued-inclusion}) is equivalent to the following

@@

 &  & \necess-\left[\frameL\left(\ket{\psi_{0}}\right),\frameR\left(\ket{\psi_{0}}\right)\right]\subseteq\left[\frameL\left(\ket{\psi_{1}}\right),\frameR\left(\ket{\psi_{1}}\right)\right]+\left[\frameL\left(\ket{\psi_{2}}\right),\frameR\left(\ket{\psi_{2}}\right)\right]\label{eq:interval-valued-frame-function}\\
 & \Leftrightarrow & \left[1-\frameR\left(\ket{\psi_{0}}\right),1-\frameL\left(\ket{\psi_{0}}\right)\right]\subseteq\left[\frameL\left(\ket{\psi_{1}}\right)+\frameL\left(\ket{\psi_{2}}\right),\frameR\left(\ket{\psi_{1}}\right)+\frameR\left(\ket{\psi_{2}}\right)\right]\,.\nonumber 

@@

Since $\left\Vert \mathscr{I}\right\Vert \le\varepsilon$, the length

@@

In another word, the left-end function~$\frameL$ satisfies the following
inequalities

@@

state~$\rho$ such that 

@@

for any state $\ket{\psi}\in\CP{2}$. Our Question~\ref{question:approximation-Gleason}

@@

With different settings, whether there is an approximate version of
Gleason's theorem was asked by Sam Sanders in constructivenews on
2013~\cite{Sanders2013}, and there is no clear answer for his question
as far as I understand. To have an idea how hard this question could
be, recall in Sec.~\ref{sec:fuzzy} we state that a quantum probability
space is glued by a family of classical probability spaces. This is
similar to the situation that a manifold is glued by many local coordinates.
When each small piece has \emph{exactly} the same and positive curvature,
the Killing-Hopf theorem asserts this manifold is a sphere, but little
can we say even if the curvature has only small deviation from constant.
Similar situation might happen when approximating Gleason's theorem.
When the sum of $\frameL$ on each ``local'' classical probability
space, $\sum_{i=0}^{2}\frameL\left(\ket{\psi_{i}}\right)$, is \emph{exactly}
the same and equal to $1$, Gleason's theorem can assert $\frameL$
can be expressed as Eq.~(\ref{eq:Gleason-for-CP2}). However, when
each local classical probability space becomes imprecise, a general
$\frameL$ might be as wild as we can imagine, and we might need to
know a little bit more, like the function is of the form $f\circ\muB_{\rho}$,
to deduce its global property.

@@

After we build the quantum interval-valued probability models, we
might think how powerful a quantum computer could be based on this
model. Since the conventional quantum circuit model manipulates the
probability amplitudes of quantum states, to build a quantum computing
model above QIVPMs, we might either consider evolving the states in
the core of a QIVPM, or find a way to evolve QIVPMs directly. Both
strategies are not straightforward. As we proved in Thm.~\ref{thm:Non-extensible-of-Gleason's},
a QIVPM might have an empty core which cannot be evolved over time.
If we want to evolve and compute QIVPMs directly, we need to get the
QIVPM of the whole system from the QIVPM for each subsystem or qubit.
However, there is no straightforward way to define the tensor product
among QIVPMs because we cannot summarize a probability measure into
a state as we did in conventional quantum theory.
